Notes and other text

### 2023-04-21 doc/manuscript/translocation.qmd
Major rewrite of Discussion text

<!--# Title should be less than 135 characters: Mark -->

@@

The splined window analysis identified 58 outlier regions for *F~ST~* and 33 outlier regions for $\pi_{diff}$ (@fig-spline-manhattan A, B).
Of these, 9 regions were outliers for both metrics and 2 of these shared regions also contained outlier SNPs.
A total of 33 annotated genes were found in the 9 shared regions (SI Dataset 2).
The largest $\pi_{diff}$, indicative of directional selection, occurred in a 163kb region on Contig19, 12.9Mb upstream of the RIN3 outlier SNP (@fig-spline-manhattan C).
This region contains approximately 500 SNPs and one annotated gene called "interferon-induced very large GTPase 1-like" (GVINP1).
Additionally, a shared outlier region on Contig1 contained two complement factor genes (C6 and C7).
Interestingly, this region had lower $\pi_{diff}$ in the recovering populations, consistent with a pattern of balancing selection.
Finally, one shared outlier region on Contig8 contained one outlier SNP (TCF19) <!--# Allie comment: Again, this is a very good reason to discuss - there are both outlier windows and individuals SNPS here making it an important part outlier region to discuss BEYOND the cool MHC connection.  -->and was within 360kb of another outlier SNP (VARS) (@fig-spline-manhattan D).
This region (854kb from the beginning of the outlier windows to the VARS SNP) contained a total of 10 annotated genes.
In *Xenopus*, four of these genes occur in the extended MHC Class I region (FLOT1, TUBB, MDC1,TCF19) and three occur in the extended MHC Class II region (HSP70, LSM2, VARS) [@ohta2006].
Therefore, we conclude that this candidate region under selection is a part of the extended MHC Class I and II complex and shows synteny with other amphibian genomes.

@@

To determine whether Bd load had a negative effect on adult survival, as would be expected if frogs were highly susceptible to Bd infection, we developed linear and multilevel models.
The best model of 1-year frog survival identified several important predictors, but Bd load at the time of translocation was not among them (bd_load; @fig-survival-postdens SI).
Important predictors included winter severity in the year following translocation (snow_t1), site elevation (elevation), and donor population (donor_72996; @fig-cond-effects).
Males had somewhat higher survival than females, but this effect was only marginally important (sex_male; @fig-cond-effects, @fig-survival-postdens SI).
Collectively, these results are consistent with frogs in recovering populations having evolved sufficient resistance/tolerance to minimize the adverse effects of Bd infection.
In addition, the moderate Bd loads that characterized frogs following translocation (@fig-bdload-beforeafter SI) and relatively high survival of frogs in numerous translocated cohorts (@fig-translocation-survival) suggests that frogs trasnslocated from recovering populations can maintain the benefits of resistance/tolerance in non-natal habitats.
Finally, in 3 locations where longer CMR time series allowed us to assess the survival of new adults recruited to the population, naturally recruited adults had equivalent or higher survival probabilities than the originally translocated adults (@fig-compare_surv_probs SI).
Therefore, frog resistance in non-natal habitats is maintained across generations.

@@

Disease-induced declines are a central challenge for the conservation of global biodiversity, but strategies to recover affected species are generally lacking.
One potential strategy, and the one we use in the current study, leverages evolutionary rescue to recover species at the landscape scale.
Although previously proposed [@mendelson2019], to our knowledge this strategy has never been successfully implemented in wildlife disease systems [but see @joseph2018 for a small-scale example].
Specifically, we provide evidence that (i) MYL frogs have experienced strong selection following exposure to a recently-emerged lethal fungal disease (Bd), and (ii) individuals from these resistant populations can be used to reestablish viable populations despite ongoing Bd infection.
These results provide a rare, and perhaps novel, example of how evolutionary rescue can be used to successfully recover a species in the presence of disease.

While previous studies have documented signatures of selection in hosts following disease-induced declines [@epstein2016, @gignoux-wolfsohn2021], a strong causal chain from selection, to an increase in host fitness in the presence of the pathogen, to subsequent host recoveries remains lacking in most wildlife-disease systems.
To robustly conclude that evolutionary rescue is the primary mechanism of host recovery, ideally evidence for each link in the causal chain should be provided.
Here, we provide evidence for each of these links.
First, our genomic analyses comparing naive and recovering MYL frog populations found notable differences in allele frequencies between adjacent Bd-naive and recovering populations, consistent with strong selection due to disease.
Importantly, some of the regions of the genome that showed signatures of selection are associated with known cellular and immunological mechanisms contributing to disease resistance in other species generally [@robledo2018; @robledo2020] and amphibians specifically [@rosenblum2008; @ellison2014; @grogan2018] (for more information, see SI X).
For example, we identified a \~940kb region on contig 8 that shows evidence of selection in recovering populations and contains eight genes associated with either the MHC Class I or Class III regions<!--# Only Class I and II are mentioned in Results. Mark: Check with Allie, the supporting information text says Class I and III -->.
These results corroborate numerous previous studies linking MHC genes to Bd survival [e.g., @savage2011; @bataille2015].
Broadly, our genomic analyses imply that selection and rapid evolutionary change has contributed to survival in MYL frogs in the presence of Bd.

Second, repeated translocations and long-term monitoring of naturally recovered populations demonstrate that fitness benefits of robust phenotypes are maintained beyond the natal environment in the presence of Bd.
Specifically, robust phenotypes in recovered populations have reduced infection intensities, below the level known to cause significant host mortality [@vredenburg2010, @knapp2016].
This phenotype persists following translocations to new environments --- translocated populations maintain infection intensities below this threshold and infection intensity is not a significant predictor of adult survival following translocation.
Moreover, high adult survival in the presence of Bd persists into subsequent generations in non-natal environments, suggesting that robust phenotypes are heritable to some degree.
Together, we show that strong selection following disease-induced declines and heritable increases in fitness point to evolutionary rescue as an important driver of observed natural recoveries and host persistence in this system [@knapp2016].
A key advance of our study is that we take these exciting results a step further, leveraging natural evolutionary rescue and translocations to promote MYL frog recovery at the landscape scale.

<!--# Comment from Allie that will be helpful in making the cases that what we are seeing is evidence of strong selection/evolutionary rescue -->

@@

-   The MYL frog system is unique in that low productivity due to short summers provide a homogenizing effect across habitat patches, potentially minimizing the effect of the local environment on landscape-level recovery compared to what we might expect in higher productivity systems (for example, tropical amphibian assemblages) <!--# Roland, I am just brainstorming here and this might not make sense. We can delete all of this if necessary. --> .
    That being said, even in this low productivity system we still observed a strong site and environmental effect on translocated adult survival.
    <!--# This is an interesting topic, but despite low productivity MYL frog habitats are highly variable: ponds, streams, meadows, alpine lakes, forested lakes, etc. I think we should drop this bullet. Mark here:  -->

@@

![For each of the naive and recovering MYL frog populations, pie charts showing allele frequencies for the stringent set of 11 outlier SNPs from 7 distinct genes: (A) LOC108802036, (B) TCF19, VARS, MAP1S, TYRP1, SRFBP1, and (C) RIN3. Charts are superimposed on a map of the Sierra Nevada study area, with Yosemite, Kings Canyon, and Sequoia National Parks (from north to south) shown in dark gray, and the range boundary for MYL frogs in this portion of the Sierra Nevada shown in light gray. The inset map locates the study area in California. <!--# Allie: Add compass rose and format distance bar to match that in Yosemite map. Relocate pie charts to avoid the perception that all recovering populations are west of naive populations. In legend, provide interpretation of "Homozygous reference", "Homozygous alternate", and "Heterozygous". How about identifying each population as R. muscosa or R. sierrae using squares and dots, respectively? The current version of the figure ignores the species identities, but including them might help with interpretation of results. Try moving inset map to lower left portion of panel - more room there to enlarge inset?  Remove distance scale from inset map. --> <!--# Bree comment: Minor but the little set-in map in the central panel may be too small to be useful. -->](figures/allele_maps.jpg){#fig-allelefrequencies}

![Results from the splined window analyses showing outlier windows for $\pi_{diff}$ (A) and *F~ST~* (B). Outlier windows for either $\pi_{diff}$ or *F~ST~* are shown in blue and outlier windows for both $\pi_{diff}$ and *F~ST~* are shown in red. (C) Magnified Contig19 from (A) showing two adjacent outlier windows for $\pi_{diff}$ 12.9Mb upstream of the RIN3 outlier SNP (indicated with a dashed vertical blue line). (D) Magnified Contig8 from from (B) showing the *F~ST~* outlier window that includes the outlier SNPs TCF19 and VARS. This region of the genome contains 10 annotated genes known to occur in the extended MHC Class I and II regions. <!--# As I commented in the Results text, numerous outlier regions were identified, so how and why were the regions highlighted in C and D chosen? Legend text that describes the outlier categories needs additional editing to clarify the categories. My attempt feels only partially successful. Capitalize plot identifiers (A, B, C, D). -->](figures/spline_manhattan_figure.jpg){#fig-spline-manhattan}

![Median 1-year survival of each cohort of translocated frogs at the 12 recipient sites included in this study. Sites are arranged along the x-axis using the average of the median 1-year survival per translocation at each site. Frog survival was estimated using CMR methods. Dot colors indicate the donor population from which frogs in each translocated cohort were collected. Error bars show the 2.5% and 97.5% posterior quantiles (i.e., 95% uncertainty intervals). When multiple translocations were conducted to a site, points and error bars are slightly offset to avoid overlap.](figures/translocation_survival_bysiteid.png){#fig-translocation-survival width="450"}

![Results from a multilevel Bayesian model showing conditional effects of the important predictors of 1-year frog survival. (A) winter severity in the year following translocation, (B) elevation of recipient site, (C) donor population, and (D) sex. In (A) and (B), blue lines are medians and gray ribbons are 95% uncertainty intervals. In (C) and (D), box plots show medians, first and third quartiles, largest and smallest values within 1.5x interquartile range, and values outside the 1.5x interquartile range.](figures/cond_effects_plot.png){#fig-cond-effects}

![A population viability analysis of translocated MYL frogs that quantifies the long-run growth rate and extinction probability of frog populations: (A) Predicted long-run growth rate $\lambda$ for different values of yearly adult survival probability $\sigma_{A_R}$ and probability of recruitment $\omega$, given the parameterized, deterministic model for translocated frogs. Colored points show the predicted $\lambda$ values for the twelve translocated populations when probability of recruitment into the adult class is $\omega = 0.3$ (given by the dashed line). The red line shows where $\lambda = 1$. (B) Predicted 50-year extinction probabilities of the 12 translocated populations given demographic stochasticity, environmental variability in $\omega$, and different mean values of $\omega$. The dashed line shows where $\omega = 0.3$, for comparison to (A). There are 6 lines at extinction probability = 1, 5 of which (70414--70619) are hidden beneath the line for 70505.](figures/pop_viability_figures_for_manuscript.jpg){#fig-viability}

![Sensitivity analysis and empirical validation of the stage-structured MYL frog model. (A) Elasticity of $\lambda$ with changes in three parameters: $\sigma_{A_R}$ (yearly survival probability of naturally recruited adults), $\omega$ (yearly probability of successful recruitment from juveniles to adults), and *F (*number of eggs produced by a female frog in a year that successfully hatch). Elasticity is calculated at the default parameter values for each population and $\omega$ = 0.3. (B) 100 simulated trajectories (black lines) that most closely matched the observed abundance trajectory of adult amphibians at site 70550 (light orange).](figures/pop_viability_figures_for_supp.jpg){#fig-viability-supp}

@@

@@

![Results from the exome capture selection study. (A) Map of sampling locations in the Sierra Nevada study area showing Bd-naive populations in red and recovering populations in blue. Yosemite, Kings Canyon, and Sequoia National Parks (from north to south) are shown in dark gray, and the range boundary for MYL frogs in this portion of the Sierra Nevada is shown in light gray. (B) PCA calculated from binary SNPs showing genomic relationship of samples. Numeric labels and colors match those in (A). (C) qqplot showing observed and expected p-values for 148,307 SNPs and indels as calculated in GEMMA. Dashed line shows Bonferroni corrected p-value for threshold of 0.05 (liberal) and 0.01 (stringent) outliers <!--# Add line for 0.05, if that distinction would be useful. -->. (D) Manhattan plot showing p-values for each SNP as calculated by GEMMA. SNPs are sorted by genomic position and contigs are sorted by size. Red line shows Bonferroni corrected p-value for threshold of 0.01. <!--# Show line for threshold of 0.05 also? -->Outlier SNPs above this threshold are highlighted and labeled. Bold labels indicate at least one non-synonymous SNP is present in that gene. <!--# Map (panel A) is duplicated here and in the mapped allele frequency figure. I think it can be removed from this figure. Capitalize panel labels. Label/identify both the liberal and stringent outlier sets? -->](figures/map_qqplot_manhattan.jpg){#fig-selectionresults}

@@

![Comparison of the average yearly adult survival probabilities for adults translocated to each of three sites versus adults that were naturally recruited at each site (as a result of reproduction by translocated frogs). In contrast to @fig-translocation-survival, these are not survival probabilities from the first year following translocation, but instead represent averaged survival probabilities across multiple years and cohort. Points are median estimates and error bars give the 95% uncertainty intervals around the estimates, accounting for yearly variation in survival probabilities. The large error bars for 70550 reflect the large within and among individual variability in yearly survival probability at this site (as also reflected in one-year survival probabilities in [@fig-translocation-survival]). All estimates are derived from the mrmr package.](figures/compare_surv_probs.jpg){#fig-compare_surv_probs}

>>>>

## Alternative title: Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

<!--# Title should be less than 135 characters -->

@@

The splined window analysis identified 58 outlier regions for *F~ST~* and 33 outlier regions for $\pi_{diff}$ (@fig-spline-manhattan A, B).
Of these, 9 regions were outliers for both metrics and 2 of these shared regions also contained outlier SNPs.
A total of 33 annotated genes were found in the 9 shared regions (SI Dataset 2).
The largest $\pi_{diff}$, indicative of directional selection, occurred in a 163kb region on Contig19, 12.9Mb upstream of the RIN3 outlier SNP (@fig-spline-manhattan C).
This region contains approximately 500 SNPs and one annotated gene called "interferon-induced very large GTPase 1-like" (GVINP1).
Additionally, a shared outlier region on Contig1 contained two complement factor genes (C6 and C7).
Interestingly, this region had lower $\pi_{diff}$ in the recovering populations, consistent with a pattern of balancing selection.
Finally, one shared outlier region on Contig8 contained one outlier SNP (TCF19) and was within 360kb of another outlier SNP (VARS) (@fig-spline-manhattan D).
This region (854kb from the beginning of the outlier windows to the VARS SNP) contained a total of 10 annotated genes.
In *Xenopus*, four of these genes occur in the extended MHC Class I region (FLOT1, TUBB, MDC1,TCF19) and four occur in the extended MHC Class III region (HSP70, LSM2, VARS <!--# Check on number of genes, add missing gene here -->) [@ohta2006].
Therefore, we conclude that this candidate region under selection is a part of the extended MHC Class I and III complex and shows synteny with other amphibian genomes.

@@

To determine whether Bd load had a negative effect on adult survival, as would be expected if frogs were highly susceptible to Bd infection, we developed linear and multilevel models.
The best model of 1-year frog survival identified several important predictors, but Bd load at the time of translocation was not among them (bd_load; @fig-survival-postdens SI).
Important predictors included winter severity in the year following translocation (snow_t1), site elevation (elevation), and donor population (donor_72996; @fig-cond-effects).
Males had somewhat higher survival than females, but this effect was only marginally important (sex_male; @fig-cond-effects, @fig-survival-postdens SI).
Collectively, these results are consistent with frogs in recovering populations having evolved sufficient resistance/tolerance to minimize the adverse effects of Bd infection.
In addition, the moderate Bd loads that characterized frogs following translocation (@fig-bdload-beforeafter SI) and relatively high survival of frogs in numerous translocated cohorts (@fig-translocation-survival) suggests that frogs translocated from recovering populations can maintain the benefits of resistance/tolerance in non-natal habitats.
Finally, in 3 locations where longer CMR time series allowed us to assess the survival of new adults recruited to the population, naturally recruited adults had equivalent or higher survival probabilities than the originally translocated adults (@fig-compare_surv_probs SI).
Therefore, frog resistance in non-natal habitats is maintained across generations.

@@

Disease-induced population declines are impacting global biodiversity [@daszak2000], and strategies to recover affected species are lacking.
Recovery in this context is likely to be particularly challenging because typically recovery will need to occur in the presence of ongoing disease.
One potential strategy, and the one we employ in the current study, utilizes populations of resistant/tolerant individuals as a source of animals with which to reestablish extirpated populations in the presence of disease.
Although previously proposed [@mendelson2019], this strategy has rarely been successfully implemented (see [@joseph2018] for a small-scale example).
In particular, although numerous studies have documented signatures of selection in hosts following disease-induced declines [e.g., @epstein2016; @gignoux-wolfsohn2021], we are unaware of any that used individuals bearing the resulting adaptive alleles to recover populations.
Here, we provide evidence that (i) MYL frogs have experienced strong selection following exposure to a recently-emerged lethal fungal disease (Bd) and that some of the genomic regions under selection are associated with genes related to disease resistance/tolerance, and (ii) reintroduction of individuals from these resistant/tolerant populations allows the successful reestablishment of viable frog populations despite ongoing Bd infection.
Although these results are generally consistent with expectations under evolutionary rescue [@carlson2014], obtaining strong evidence in wild animal populations for each of the associated causal links (i.e., selection for adaptive alleles $\rightarrow$ increase in host fitness in the presence of the pathogen $\rightarrow$ positive population growth and recovery) is difficult and perhaps impossible.
Nonetheless, in the following summary of our results, we use evolutionary rescue as an informative, but tentative framework.

Previous field studies in MYL frogs show that frog-Bd dynamics and frog survival in the presence of Bd are fundamentally different between naive and recovering populations.
Following the arrival and establishment of Bd in previously-naive populations, adult frogs develop high Bd loads that lead to mass die-offs [@vredenburg2010].
In contrast, in recovering populations adult frogs typically have low-to-moderate and relatively constant Bd loads and mass die-offs are not observed [@briggs2010; @knapp2011, see also @fig-bdload-beforeafter - before translocation].
These differences in Bd load of frogs from naive and recovering populations are also observed in controlled laboratory studies [@knapp2016, Figure 4].
The observed reduced susceptibility of frogs from recovering populations could in theory be due to several factors including natural selection for more resistant/tolerant genotypes and/or acquired immunity, but until now evidence to evaluate the role of evolution was lacking.

Results from our genomic analyses strongly suggest that natural selection for adaptive alleles and resulting increased resistance/tolerance is at least partially responsible for the reduced susceptibility of frogs in recovering populations.
We identified substantial differences in allele frequencies and larger genomic regions between adjacent Bd-naive and recovering MYL frog populations, consistent with strong selection following Bd exposure.
Importantly, some regions under selection are associated with cellular and immunological mechanisms known to contribute to disease resistance, including in amphibians [@grogan2018, reviewed in @zamudio2020].
For example, the major histocompatibility complex (MHC) plays an important role in immunity. In our study, we identified a region that shows evidence of selection in recovering populations and contains eight genes associated with either the MHC Class I or Class III regions. These results corroborate numerous previous studies linking MHC genes to amphibian resistance/tolerance against Bd \[e.g., @savage2011; @bataille2015\].
Similarly, the region with the strongest indication of directional selection contains the gene, GVINP1.
Several previous studies have found this gene to be differentially expressed during Bd infection [e.g., @grogan2018; @ellison2014].
This gene was also strongly linked to disease in salmon, explaining a remarkable 20% of the resistance phenotype [@robledo2020; @robledo2018].
Finally, the RIN3 gene showed a consistent pattern of allele frequency differences across all nine of the frog populations sampled in this study, indicating consistent selection across a wide geographic area.
We provide additional details on these and other genes identified in our study as under selection in the SI.

Reintroductions are an increasingly used method to address escalating loss of wildlife species and populations.
However, recent reviews suggest that such efforts generality have a low success rate (Griffiths Pavajeau 2008, Dodd 2005).
Of the 24 translocations of MYL frogs that we conducted as part of the current study, x% showed evidence of successful reproduction, x% of recruitment of new adults, and x% showed high annual survival and a high probability of persistence over 50 years.
Following translocation, Bd loads were relatively unchanged.
In addition, survival/, low-moderate Bd loads extend beyond the natal environment, maintained between generations.
Success was even higher for more recently conducted translocations.
The xx% success rate for these more recent translocations is particularly remarkable given that population establishment is occurring in the presence of Bd, a pathogen to which MYL frogs are highly susceptible.

several lines of evidence indicate that the selection-derived resistance/tolerance and increased fitness in the presence of Bd is associated with population recovery.
Results from frog translocations conducted as part of the current study indicate that following translocation Bd loads in frogs changed little and post-translocation survival was relatively high at most sites.
These results are in contrast to the outcome of translocations we conducted previously of naive frogs into Bd-positive habitats (conducted prior to our current understanding of the importance of Bd exposure history on frog susceptibilty).
Following those translocations, frogs developed high Bd loads and experienced high mortality (details in SI <!--# These data are unpublished, so if we want to refer to them we'll have to include the data in the SI. Is this necessary to support this result? I don't see any obvious alternatives. -->).
In addition, in the current translocations increased fitness maintained between generations.

Results from frog translocations conducted as part of the current study indicate that higher fitness in the presence of Bd observed in frogs from recovering populations extends beyond the natal environment.
Specifically, following translocation of frogs from natal to recipient sites, Finally, in translocated populations the higher fitness is also maintained in subsequent generations (@fig-compare_surv_probs).
Collectively, these results indicate that the higher fitness observed in recovering Bd-positive populations is not simply a site effect, but also a frog effects and not only the sites, suggesting that the observed resistance and resulting h

These are that the increased fitness repeated translocations and long-term assessment of naturally-recovering populations demonstrate that fitness benefits of robust phenotypes are maintained beyond the natal environment in the presence of Bd.
Specifically, robust phenotypes in recovered populations have reduced Bd loads, below the level known to cause significant host mortality [@vredenburg2010, @knapp2016].
This phenotype persists following translocations to new environments --- translocated populations maintain infection intensities below this threshold and infection intensity is not a significant predictor of adult survival following translocation.
Moreover, high adult survival in the presence of Bd persists into subsequent generations in non-natal environments, suggesting that robust phenotypes are heritable to some degree.
Together, we show that strong selection following disease-induced declines and heritable increases in fitness point to evolutionary rescue as an important driver of observed natural recoveries and host persistence in this system [@knapp2016].
A key advance of our study is that we take these exciting results a step further, leveraging natural evolutionary rescue and translocations to promote MYL frog recovery at the landscape scale.

<!--# Comment from Allie that will be helpful in making the case that what we are seeing is evidence of strong selection/evolutionary rescue -->

@@

@@

![For each of the naive and recovering MYL frog populations, pie charts showing allele frequencies for the stringent set of 11 outlier SNPs from 7 distinct genes: (A) LOC108802036, (B) TCF19, VARS, MAP1S, TYRP1, SRFBP1, and (C) RIN3. Charts are superimposed on a map of the Sierra Nevada study area, with Yosemite, Kings Canyon, and Sequoia National Parks (from north to south) shown in dark gray, and the range boundary for MYL frogs in this portion of the Sierra Nevada shown in light gray. The inset map locates the study area in California. <!--# Allie: Add compass rose and format distance bar to match that in Yosemite map. Relocate pie charts to avoid the perception that all recovering populations are west of naive populations. In legend, provide interpretation of "Homozygous reference", "Homozygous alternate", and "Heterozygous". How about identifying each population as R. muscosa or R. sierrae using squares and dots, respectively? The current version of the figure ignores the species identities, but including them might help with interpretation of results. Try moving inset map to lower left portion of panel - more room there to enlarge inset?  Remove distance scale from inset map. --> <!--# Bree comment: Minor but the little set-in map in the central panel may be too small to be useful. -->](figures/allele_maps.png){#fig-allelefrequencies width="450"}

![Results from the splined window analyses showing outlier windows for $\pi_{diff}$ (A) and *F~ST~* (B). Outlier windows for either $\pi_{diff}$ or *F~ST~* are shown in blue and outlier windows for both $\pi_{diff}$ and *F~ST~* are shown in red. (C) Magnified Contig19 from (A) showing two adjacent outlier windows for $\pi_{diff}$ 12.9Mb upstream of the RIN3 outlier SNP (indicated with a dashed vertical blue line). (D) Magnified Contig8 from from (B) showing the *F~ST~* outlier window that includes the outlier SNPs TCF19 and VARS. This region of the genome contains 10 annotated genes known to occur in the extended MHC Class I and II regions. <!--# As I commented in the Results text, numerous outlier regions were identified, so how and why were the regions highlighted in C and D chosen? Legend text that describes the outlier categories needs additional editing to clarify the categories. My attempt feels only partially successful. Capitalize plot identifiers (A, B, C, D). -->](figures/splinewindow_manhattan.png){#fig-spline-manhattan width="450"}

![Median 1-year survival of each cohort of translocated frogs at the 12 recipient sites included in this study. Sites are arranged along the x-axis using the average of the median 1-year survival per translocation at each site. Frog survival was estimated using CMR methods. Dot colors indicate the donor population from which frogs in each translocated cohort were collected. Error bars show the 2.5% and 97.5% posterior quantiles (i.e., 95% uncertainty intervals). When multiple translocations were conducted to a site, points and error bars are slightly offset to avoid overlap.](figures/translocation_survival_bysiteid.png){#fig-translocation-survival width="300"}

![Results from a multilevel Bayesian model showing conditional effects of the important predictors of 1-year frog survival. (A) winter severity in the year following translocation, (B) elevation of recipient site, (C) donor population, and (D) sex. In (A) and (B), blue lines are medians and gray ribbons are 95% uncertainty intervals. In (C) and (D), box plots show medians, first and third quartiles, largest and smallest values within 1.5x interquartile range, and values outside the 1.5x interquartile range.](figures/cond_effects_plot.png){#fig-cond-effects width="450"}

![A population viability analysis of translocated MYL frogs that quantifies the long-run growth rate and extinction probability of frog populations: (A) Predicted long-run growth rate $\lambda$ for different values of yearly adult survival probability $\sigma_{A_R}$ and probability of recruitment $\omega$, given the parameterized, deterministic model for translocated frogs. Colored points show the predicted $\lambda$ values for the twelve translocated populations when probability of recruitment into the adult class is $\omega = 0.3$ (given by the dashed line). The red line shows where $\lambda = 1$. (B) Predicted 50-year extinction probabilities of the 12 translocated populations given demographic stochasticity, environmental variability in $\omega$, and different mean values of $\omega$. The dashed line shows where $\omega = 0.3$, for comparison to (A). There are 6 lines at extinction probability = 1, 5 of which (70414--70619) are hidden beneath the line for 70505.](figures/pop_viability_figures_for_manuscript1.jpg){#fig-viability}

![Sensitivity analysis and empirical validation of the stage-structured MYL frog model. (A) Elasticity of $\lambda$ with changes in three parameters: $\sigma_{A_R}$ (yearly survival probability of naturally recruited adults), $\omega$ (yearly probability of successful recruitment from juveniles to adults), and *F (*number of eggs produced by a female frog in a year that successfully hatch). Elasticity is calculated at the default parameter values for each population and $\omega$ = 0.3. (B) 100 simulated trajectories (black lines) that most closely matched the observed abundance trajectory of adult amphibians at site 70550 (light orange).](figures/pop_viability_figures_for_supp1.jpg){#fig-viability-supp}

@@

### Translocations

<!--# Necessary? -->

Description of translocation of Bd-naive populations to Bd-positive sites

-   Spur Basin to Pinchot, Upper, and Ruskin Basins

-   Marmot to Square Lake

Description of translocation of Bd-naive populations to Bd-negative sites

-   Marmot to Ameletus Lake

-   Upper Center Basin to South Center Basin

@@

![Results from the exome capture selection study. (A) Map of sampling locations in the Sierra Nevada study area showing Bd-naive populations in red and recovering populations in blue. Yosemite, Kings Canyon, and Sequoia National Parks (from north to south) are shown in dark gray, and the range boundary for MYL frogs in this portion of the Sierra Nevada is shown in light gray. (B) PCA calculated from binary SNPs showing genomic relationship of samples. Numeric labels and colors match those in (A). (C) qqplot showing observed and expected p-values for 148,307 SNPs and indels as calculated in GEMMA. Dashed line shows Bonferroni corrected p-value for threshold of 0.05 (liberal) and 0.01 (stringent) outliers <!--# Add line for 0.05, if that distinction would be useful. -->. (D) Manhattan plot showing p-values for each SNP as calculated by GEMMA. SNPs are sorted by genomic position and contigs are sorted by size. Red line shows Bonferroni corrected p-value for threshold of 0.01. <!--# Show line for threshold of 0.05 also? -->Outlier SNPs above this threshold are highlighted and labeled. Bold labels indicate at least one non-synonymous SNP is present in that gene. <!--# Map (panel A) is duplicated here and in the mapped allele frequency figure. I think it can be removed from this figure. Capitalize panel labels. Label/identify both the liberal and stringent outlier sets? -->](figures/pca_qq_manhattan.png){#fig-selectionresults width="450"}

@@

![Comparison of the average yearly adult survival probabilities for adults translocated to each of three sites versus adults that were naturally recruited at each site (as a result of reproduction by translocated frogs). In contrast to @fig-translocation-survival, these are not survival probabilities from the first year following translocation, but instead represent averaged survival probabilities across multiple years and cohort. Points are median estimates and error bars give the 95% uncertainty intervals around the estimates, accounting for yearly variation in survival probabilities. The large error bars for 70550 reflect the large within and among individual variability in yearly survival probability at this site (as also reflected in one-year survival probabilities in [@fig-translocation-survival]). All estimates are derived from the mrmr package.](figures/compare_surv_probs.jpg){#fig-compare_surv_probs width="450"}
